\speculake: Hiding Malware in Speculative Execution

### 2018-12-10 paper/architecture.tex
fix typos add clarifications

In this subsection, we will describe the trigger program variant, and defer discussion 
on how input data can be used a speculative trigger to
Section~\ref{subsec:spec-buffer-overflow}.

@@

In our payload program, we first perform the same indirect jumps. We ensure
the source and destination addresses of these jumps is the same as in the
trigger program by manually defining their containing function at a fixed
address inside a linker script. We also do the final indirect call to a function
pointer, but with two differences. First, the destination in the function
pointer is a different address, and second, the memory location of the function
pointer itself is uncached. This forces the CPU to predict the destination of
the final indirect call while it waits for the function pointer to load from
memory. Due to the similar history of branches with the trigger program, the CPU
will (incorrectly) predict the destination to be the same as the one in the
trigger program, which determines the speculative entry point for the payload.
Even though the in-order execution of payload program never executes or even
reads from this address, the CPU will briefly execute instructions there
speculatively. \processorTable

In Table~\ref{table:processors} we analyze the number of necessary training
indirect jumps various processors require to consistently ($> 95\%$) have the
payload program enter the speculative world at the location primed in the
trigger program. We found that 28 indirect jumps was sufficient for our trigger
program on each of the test processors to reliably ensure the speculative
execution began at the correct address.

@@

We find that when executing speculatively, the number of instructions completed
is rarely limited by cache miss duration. Instead
ROB size and processor mechanisms for resolving
data dependencies define an upper boundary of approximately 150
instructions\footnote{While \texttt{nop} is able to execute up to the full 220
ROB capacity, instructions that do useful work (and/or use multiple $\mu$ ops) cannot reach this limit}.

>>>>

In this subsection, we will describe the trigger program variant, and defer
discussion on how input data can be used as a speculative trigger to
Section~\ref{subsec:spec-buffer-overflow}.

@@

In our payload program, we first perform the same indirect jumps. We ensure
the source and destination addresses of these jumps is the same as in the
trigger program by manually defining their containing function at a fixed
address inside a linker script. We also do the final indirect call to a function
pointer, but with two differences. First, the destination in the function
pointer is a different address, and second, the memory location of the function
pointer itself is uncached. This forces the CPU to predict the destination of
the final indirect call while it waits for the function pointer to load from
memory. Due to the similar history of branches with the trigger program, the CPU
will (incorrectly) predict the destination to be the same as the one in the
trigger program, which determines the speculative entry point for the payload.
Even though the in-order execution of payload program never executes or even
reads from this address, the CPU will briefly execute instructions there
speculatively.

\processorTable

In Table~\ref{table:processors} we analyze the number of necessary training
indirect jumps various processors require to consistently ($> 95\%$) have the
payload program enter the speculative world at the chosen speculative entry
point in the trigger program. We found that 28 indirect jumps was sufficient for
our trigger program on each of the test processors to reliably ensure the
speculative execution began at the correct speculative entry point.

@@

We find that when executing speculatively, the number of instructions completed
is rarely limited by cache miss duration. While data dependency and ROB size
have an effect on the number of instructions that can be executed speculatively
Figure~\ref{fig:spec-capacity-np} demonstrates that the number of $\mu$-ops of
the instructions is the major limiting factor that define an upper boundary of
approximately 150 instructions\footnote{While \texttt{nop} is able to execute up
to the full 220 ROB capacity, instructions that do useful work (and/or use
multiple $\mu$ ops) cannot reach this limit}.

### 2018-12-10 paper/figures.tex
fix typos add clarifications

\newcommand{\processorTable}{
\begin{table}
    \centering
    % \resizebox{\linewidth}{!}{\begin{tabularx}{\linewidth}{X X X| Y Y Y}
    \begin{tabularx}{\linewidth}{X b b| Y Y Y}
        \ \newline
        \ \newline
        Processor & \ \newline
        Year\newline
        released & \ \newline
        Micro\newline
        arch & \ \newline
        Nested\newline 
        Spec? & Num\newline
        indirect jumps* & Max $\mu$~ops\newline
        (nop)\\
        \hline
        Intel Xeon CPU E3-1270 v6 & 2017 & Skylake & \checkmark & 28 & 220\\
        Intel Core i5-7200U       & 2016 & Skylake & \checkmark & 26 & 220\\
        Intel Xeon CPU E3-1276 v3 & 2014 & Haswell &            & 26 & 178\\
        AMD EPYC 7251             & 2017 & Ryzen   &            & 4  & 178\\
        \\
    \end{tabularx}%}
    \caption{\textbf{Processor features}--- We analyzed the capability of
    \speculake on three Intel and one AMD processor. Both Skylake processors
    were capable of nested speculation (Section~\ref{sssec:nested-spec}) though
    the older Haswell processor and the AMD processor were not. We found that
    the Intel Xeon CPU E3-1270 v6 required 28 identical indirect jumps in the
    trigger and payload programs to reliably enter the speculative world at the
    desired address ($> 95\%$) while the Intel Core i5-7200U only required 26
    jumps.}
    \label{table:processors}
\end{table}
}

>>>>

\newcommand{\processorTable}{
\begin{table}
    \centering
    % \resizebox{\linewidth}{!}{\begin{tabularx}{\linewidth}{X X X| Y Y Y}
    \begin{tabularx}{\linewidth}{X b b| Y Y Y}
        \ \newline
        \ \newline
        Processor & \ \newline
        Year\newline
        released & \ \newline
        Micro\newline
        arch & \ \newline
        Nested\newline 
        Spec? & Num\newline
        indirect jumps* & Max $\mu$~ops\newline
        (nop)\\
        \hline
        Intel Xeon CPU E3-1270 v6 & 2017 & Skylake & \checkmark & 28 & 220\\
        Intel Core i5-7200U       & 2016 & Skylake & \checkmark & 26 & 220\\
        Intel Xeon CPU E3-1276 v3 & 2014 & Haswell &            & 26 & 178\\
        AMD EPYC 7251             & 2017 & Ryzen   &            & 4  & 178\\
        \\
    \end{tabularx}%}
    \caption{\textbf{Processor features}--- We analyzed the capability of
    \speculake on one AMD and three Intel processors. Both Skylake processors
    were capable of nested speculation (Section~\ref{sssec:nested-spec}) though
    the older Haswell processor and the AMD processor were not. We found that
    the Intel Xeon CPU E3-1270 v6 required 28 identical indirect jumps in the
    trigger and payload programs to reliably enter the speculative world at the
    desired speculative entry point ($> 95\%$) while the Intel Core i5-7200U
    and Intel Xeon CPU E3-1276 v3 only required 26 jumps. The AMD EPYC 7251
    processor maintained a 85\% speculative hit rate with as low as 4 and as
    high as 32 training jumps.}
    \label{table:processors}
\end{table}
}

### 2018-12-10 paper/payloads.tex
fix typos add clarifications

For example, if the trigger program makes 30 training jumps, followed by 10
additional indirect jumps, and the payload program performs the same 30 training
jumps before a stall, the CPU will predict the payload program will also perform
the next 10 jumps. Speculatively the CPU will attempt to follow each of these
jumps speculatively.

If each jump has the potential to land in up to 1024 possible locations, this
would provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of
security. We describe our implementation of nested speculative execution in
Section~\ref{subsec:impl-aes}, where we achieve 128~bits of security.
Figure~\ref{fig:nested-spec} demonstrates this in practice, the trigger program
running a series of training jumps followed by indirect jumps influences the
payload program to follow the same path. Any analyst attempting to follow this
path without the trigger program would need to correctly jump to each of the 10
addresses to load the correct key. The security gained here is not a result of a
single entry point remaining secret but rather multiple independent jumps by the
unknown trigger program providing numerous landing points for each jump.

>>>>

For example, if the trigger program makes 30 training jumps, followed by 10
additional indirect jumps, and the payload program performs the same 30 training
jumps before a stall, the CPU will predict the payload program will also perform
the next 10 jumps. Speculatively the CPU will attempt to follow each of these
jumps.

If each jump has the potential to land in up to 1024 possible locations, this
would provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of
security. We describe our implementation of nested speculative execution in
Section~\ref{subsec:impl-aes}, where we achieve 128~bits of security.
Figure~\ref{fig:nested-spec} demonstrates this in practice, the trigger program
running a series of training jumps followed by indirect jumps influences the
payload program to follow the same path. Any analyst attempting to follow this
path without the trigger program would need to correctly jump to each of the 10
addresses to load the correct key. The secret key created here is not a result
of a single entry point remaining secret (via the branch predictor) but rather
multiple independent jumps by the unknown trigger program providing numerous
potential speculative entry points for each jump.
